---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# shortIRT

<!-- badges: start -->
<!-- badges: end -->

The goal of shortIRT is to simple tool for the development of static short test forms (STFs) in an Item Response Theory (IRT) based framework. Specifically, two main procedures are considered: 

1. The typical IRT-based procedure for the development of STFs (here denoted as benchmark procedure, BP), according to which the most informative items are selected without considering any specific level of the latent trait

2. The IRT procedure based on the definition of levels of interest of the latent trait (i.e., $\theta$ targets, here denoted as $\theta$-target procedure). The selected items are those most informative in respect to the $\theta$ targets. This procedure can be further categorized according to the methodology used for the definition of the $\theta$ targets: 
    a. Equal interval procedure (EIP): The latent trait is divided in $n + 1$ (where $n$ is the number of items to be included in the STF) intervals of equal width and the central points of each interval are the $\theta$ targets
    b. Unequal interval procedure (UIP): The latent trait is clustered in $n$ clusters (where $n$ is the number of items to be included in the STF) and centroids of each clusters are the $\theta$ targets
    c. User defined procedure (UDP): The user manually defines the $\theta$ targets to which the STF should tend to. They migth also be the same $\theta$ values (e.g., for the development of a screening STF with a cut-off point).

## Installation

You can install the development version of shortIRT from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("OttaviaE/shortIRT")
```

## Example

In the example, a 5-item STF is developed from a full-length test of 20 items.

```{r example}
library(shortIRT)
```

First, generate the responses of 1,000 respondents to the 20 items of the full-length test:  


```{r cars}
 # Simulate person and item parameters
 true_theta <- rnorm(1000)
 b <- runif(30, -3, 3)
 a <- runif(30, 0.6, 2)
 parameters <- data.frame(b, a)
 # simulate data
 data <- sirt::sim.raschtype(true_theta, b = b, fixed.a = a)
```

Then, use the EIP for the development of the 5-item STF, incuding also the matrix of the item parameters and the true $\theta$ of the subjects (these arguments are not mandatory).

```{r}
stf_eip <- eip(data, item_par = parameters, starting_theta = true_theta, num_item = 5)
# Show the items included in the STF, the theta targets, and the IIFs of 
# the optimal item for each theta target
stf$item_stf
```

Plot the test information function of both the full-length test and the STF: 

```{r put.width = "100%", fig.align='center'}
plot_tif(stf_eip, tif = "both")  + theme_light() + 
  ylab("Information") + xlab(expression(theta)) + 
  theme(axis.title = element_text(size = 26), 
        strip.text.x = element_text(size = 28))
```

Now, the bias between the true $\theta$ and the estimated $\theta$ ($\hat{\theta}$) can be computed as well with the `diff_theta()` function, which returns both the difference and the absolute difference between $\theta$s: 
```{r}
bias_eip <- diff_theta(stf_eip, true_theta = true_theta)
head(bias_eip)
```

Finally, bias can also be plotted with the function `plot_difference()` as a function of different levels of $\theta$. Default is to plot the difference as a function of 4 levels of $\theta$:

```{r}
plot_difference(bias_eip) + theme_light() + 
  ylab("Difference") + xlab(expression(theta)) + 
  theme(axis.title = element_text(size = 26), 
        strip.text.x = element_text(size = 28), 
        axis.text = element_text(size = 22))
```

But this can be changed by changing the default of the arguments `type` and `levels` to plot, for instance, the absolute difference for 20 levels of the latent trait: 

```{r}
plot_difference(bias_eip, type = "absolute_diff", levels = 20) + theme_light()  + xlab(expression(theta)) + 
  theme(axis.title = element_text(size = 26), 
        strip.text.x = element_text(size = 28), 
        axis.text = element_text(size = 18, angle = 45, hjust=1))
```
